---
title: "Authoring and validating question banks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Authoring and validating question banks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## The index schema

The index is a UTF-8 CSV with one row per question and seven required columns:

| Column | Meaning |
|---|---|
| `id` | Stable, unique identifier used for selection and variable prefixes |
| `file` | Absolute path or, preferably, a path relative to the index |
| `title` | Human-readable catalogue title |
| `topic` | Filterable topic label |
| `difficulty` | Filterable difficulty label such as `easy` or `hard` |
| `marks` | Positive numerical mark value |
| `type` | One of `numerical`, `shortanswer`, `singlechoice`, or `multichoice` |

Additional columns are allowed and are carried into
`selected-questions.csv`. This makes it possible to add features such as week,
learning outcome, source dataset, author, or review status without changing the
package.

Relative paths make a bank portable. For example:

```csv
id,file,title,topic,difficulty,marks,type
joins-01,questions/joins/joins-01.qmd,Choose a join,Joins,easy,1,singlechoice
```

## The fragment contract

Each `.qmd` or `.Rmd` fragment must:

1. contain no YAML front matter;
2. have exactly one H2 (`##`) question heading;
3. have exactly one complete `::: unilur-solution` block;
4. use R chunk labels that are unique across the entire bank;
5. prefix question variables with `q_<question_id>_`, replacing hyphens in the
   ID with underscores; and
6. call the helper corresponding to the indexed `type`.

A numerical fragment looks like this:

````markdown
## Calculate a total

```{r arithmetic-01-calculate, echo=FALSE}
q_arithmetic_01_answer <- 6 * 7
```

What is six multiplied by seven?

**Answer:** `r paste0(intToUtf8(96), "r quiz_numeric(q_arithmetic_01_answer)", intToUtf8(96))`

::: unilur-solution
Multiplication gives **`r paste0(intToUtf8(96), "r q_arithmetic_01_answer", intToUtf8(96))`**.
:::
````

Do not put YAML in fragments. `build_quiz()` creates complete wrapper documents
and supplies their YAML metadata.

## Answer helpers

Use exactly one of these families for the indexed type:

```{r eval = FALSE}
quiz_numeric(answer = 42, tolerance = 0)
quiz_shortanswer(answer = "left_join")
quiz_singlechoice(
  answer = "left_join",
  options = c("inner_join", "left_join", "anti_join")
)
quiz_multichoice(
  answer = c("filter", "select"),
  options = c("filter", "select", "lm", "hist")
)
```

The helpers adapt to the output target. Student HTML receives empty input
controls, teacher HTML marks correct responses, PDFs receive printable fields
or symbols, and Moodle receives cloze grading syntax. Do not call
`moodlequiz::cloze()` directly in a fragment, because doing so would make the
student and teacher versions inconsistent.

For single-choice questions, `answer` should equal exactly one entry in
`options`. For multiple-choice questions, every answer should appear in
`options`. Keep options as character vectors so that labels render predictably.

## Add a question step by step

1. Choose a stable ID. Avoid renaming it after quizzes have used it.
2. Add a `.qmd` fragment below `questions/`.
3. Give every R chunk a bank-wide unique label.
4. Prefix temporary variables using the ID rule.
5. Use the matching answer helper.
6. Add a worked solution block.
7. Add the index row, including its relative file path and metadata.
8. Run `validate_question_bank(index_file)`.
9. Run a quick source build with `render = FALSE`.
10. Render and review both student and teacher versions before publishing.

## What validation catches

`validate_question_bank()` checks the full catalogue for missing or blank
metadata, duplicate IDs, invalid marks, unsupported types, missing files,
front matter in fragments, missing or repeated H2 headings and solution blocks,
duplicate chunk labels, incorrect variable prefixes, direct cloze calls, and a
mismatch between the indexed type and answer helper.

Validation is structural. It cannot decide whether a question is pedagogically
clear or whether its answer is factually correct. Peer review and rendered
teacher-version review remain essential.

## Version-control recommendations

Commit the index and fragments. Usually ignore generated wrappers and rendered
outputs, or publish reviewed outputs as release artifacts. Record the version
of any data package or external source used to calculate answers. Review index
changes as carefully as code because changing marks, type, or file mappings can
change an assessment.
