---
title: "Getting started with ecosanctuary"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with ecosanctuary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## What belongs in the package?

`ecosanctuary` owns the reusable mechanics: reading metadata, validating
question fragments, selecting questions, producing wrappers, formatting answer
controls, and rendering outputs. A question-bank project owns the teaching
content: its index, question fragments, source notes, and generated products.

Keeping these responsibilities separate means that one installed package can
build many banks. Fixes to validation and rendering are made once in the
package, rather than copied into every course repository.

A bank normally looks like this:

```
my-question-bank/
├── question-index.csv
├── questions/
│   ├── easy/
│   │   └── arithmetic-01.qmd
│   └── medium/
│       └── joins-01.qmd
├── build-quiz.R
├── generated/             # created by build_quiz()
└── output/                # created by build_quiz()
```

## Why the build asks for an index file

The index file is the better primary input. A questions folder can tell the
package which files exist, but it cannot reliably supply stable IDs, display
titles, topics, difficulty, marks, or Moodle answer types. Encoding that
information only in directory and file names makes reorganising a bank risky.

The CSV is an explicit, reviewable contract. Each row connects a stable ID and
metadata to one fragment. Paths are interpreted relative to the index, so the
whole bank remains portable. The package still supports a folder-first start:
[create_question_index()] scans existing fragments and writes a template that
you complete once.

## Install the package

For ordinary use, install from GitHub:

```{r eval = FALSE}
remotes::install_github("vahdatjavad/ecosanctuary")
```

During package development, open `ecosanctuary.Rproj` and use:

```{r eval = FALSE}
devtools::document()
devtools::test()
devtools::check()
devtools::install()
```

## Create or locate an index

For an existing bank, keep the path explicit:

```{r eval = FALSE}
index_file <- "/path/to/my-question-bank/question-index.csv"
```

To bootstrap a bank from a folder:

```{r eval = FALSE}
create_question_index(
  questions_dir = "/path/to/my-question-bank/questions",
  output_file = "/path/to/my-question-bank/question-index.csv"
)
```

Open the CSV and complete every blank metadata field. Then run:

```{r eval = FALSE}
library(ecosanctuary)
validate_question_bank(index_file)
```

Validation is intentionally performed on the complete bank before any subset
is selected. A broken unselected question therefore cannot remain hidden.

## Build all or selected questions

Build all questions in catalogue order:

```{r eval = FALSE}
build_quiz(index_file = index_file)
```

Build an explicit ordered selection:

```{r eval = FALSE}
build_quiz(
  index_file = index_file,
  ids = c("joins-03", "exploration-01")
)
```

The order in `ids` becomes the question order in every output. For a random
quiz, use `n` and `seed`; do not supply `ids` at the same time.

During authoring, set `render = FALSE` for a fast structural build. This writes
the wrappers and selection manifest but does not invoke Quarto, Pandoc, Moodle,
or LaTeX.

```{r eval = FALSE}
build_quiz(index_file = index_file, render = FALSE)
```

## Use the interactive build script

Copy `system.file("examples", "build-quiz.R", package = "ecosanctuary")` into
your bank and edit its configuration. The repository-level `build-quiz.R` also
demonstrates a file chooser: when sourced interactively, it asks for the index
CSV. In non-interactive environments, set `ECOSANCTUARY_QUESTION_INDEX`.

Continue with `vignette("authoring-question-banks")` before adding questions,
then read `vignette("rendering-and-moodle")` for build dependencies and import
steps.
