---
title: "Rendering quizzes and importing Moodle XML"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rendering quizzes and importing Moodle XML}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Rendering requirements

Install the optional R packages used by a complete build:

```{r eval = FALSE}
install.packages(c("moodlequiz", "quarto", "rmarkdown"))
```

HTML and PDF builds need the Quarto command-line application. PDF output also
needs a LaTeX distribution; TinyTeX is a convenient option:

```{r eval = FALSE}
install.packages("tinytex")
tinytex::install_tinytex()
```

These tools are not required for reading, validating, or selecting questions,
or for `build_quiz(..., render = FALSE)`.

## Select questions

Use explicit IDs for a curated assessment:

```{r eval = FALSE}
build_quiz(
  index_file = index_file,
  ids = c("easy-01", "medium-04", "hard-02")
)
```

The given order is retained. Duplicate or unknown IDs are rejected.

Use sampling for practice versions or variants:

```{r eval = FALSE}
build_quiz(
  index_file = index_file,
  n = 12,
  seed = 2026,
  topics = c("Wrangling", "Visualisation"),
  difficulties = c("easy", "medium"),
  shuffle = FALSE
)
```

Sampling is without replacement and stratified by topic. The allocation is as
even as topic capacities allow; if one topic has too few questions, its unused
slots are redistributed across the other topics. A local seed makes selection
reproducible without changing the caller's random-number state. Filters limit
the pool before sampling.

With `shuffle = FALSE` (the default), topics appear in the order supplied to
`topics`, and questions within each topic retain index order. If `topics` is
omitted, all selected questions retain index order. Set `shuffle = TRUE` to
randomise the final question order.

## Build products

The default build writes:

```
generated/
├── _quarto.yml
├── quiz-moodle.Rmd
├── quiz-student.qmd
└── quiz-teacher.qmd

output/
├── quiz-moodle.xml
├── quiz-student.html
├── quiz-student.pdf
├── quiz-teacher.html
├── quiz-teacher.pdf
└── selected-questions.csv
```

The student documents omit solution blocks and show blank answer controls. The
teacher documents retain worked solutions and display correct answers. The
HTML controls are for local practice only; they do not submit or store answers.
The Moodle XML is the gradable artifact.

When `generated/` does not already contain a `_quarto.yml`, the build creates
one that makes it an isolated default Quarto project. This prevents a
surrounding website, book, or other parent Quarto project from contributing its
files or resources to the quiz output. An existing `_quarto.yml` is preserved.

Student and teacher HTML/PDF documents include a table of contents by default.
To omit it from their generated `.qmd` wrappers and rendered outputs, use:

```{r eval = FALSE}
build_quiz(index_file = index_file, toc = FALSE)
```

The manifest records question number, ID, path, title, topic, difficulty,
marks, type, and any extra index columns. Keep it with a delivered assessment
so the selection can be audited later.

## Moodle import

The Moodle interface varies by version and local configuration, but the usual
workflow is:

1. Build the quiz and confirm that `quiz-moodle.xml` exists.
2. Review `quiz-teacher.html` and the selection manifest.
3. In Moodle's question bank, choose the import action and Moodle XML format.
4. Upload `quiz-moodle.xml` and import it into the intended category.
5. Inspect the imported questions and preview grading before adding them to a
   live quiz.

The package checks that the produced XML contains the expected number of cloze
questions and no unexpected description-only questions. This is a useful
structural safeguard, not a replacement for Moodle preview and teaching review.

## Replicates and categories

Pass `category` to choose the category metadata sent to `moodlequiz`. Pass a
positive integer `replicates` to request generated variants supported by that
package:

```{r eval = FALSE}
build_quiz(
  index_file = index_file,
  category = "STAT101 / Week 4",
  replicates = 3
)
```

If question code uses randomness, set or derive question-level seeds inside the
fragment when exact reproducibility is required. The `seed` argument to
`build_quiz()` controls question selection, not calculations inside questions.

## Troubleshooting

- **An indexed file is missing:** resolve the `file` value relative to the
  index location and prefer forward slashes in committed CSV files.
- **A helper-type mismatch is reported:** use `quiz_numeric()` for `numerical`;
  the other type names match their helper suffixes.
- **Quarto cannot be found:** install the Quarto application and confirm that
  `quarto::quarto_path()` returns a path.
- **PDF rendering fails:** confirm LaTeX is installed by running
  `tinytex::is_tinytex()` or checking your system TeX installation.
- **A build works only on one computer:** remove machine-specific absolute paths
  from the index and use paths relative to the index CSV.
